---
title: "Cars"
author: "LIANG"
date: "3/30/2022"
output: 
  html_document:
    toc: yes
    toc_float: yes
    toc_depth: 6
    df_print: paged
    theme: paper
    number_of_section: yes
    hightlight: kate
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=F, error=F, messages=F)
carsDF <- read.csv("/Users/LYQ/Desktop/cars.csv")
carsDF
library(dplyr)
library(ggplot2)
```

## **Import Dataset**
Import Cars.csv into carsDF variable
```{r cars}
carsDF <- read.csv("/Users/LYQ/Desktop/cars.csv")
carsDF
```


## **First 5 cars**
Question: With tidyverse approach, show the name of the first 5 cars by the number of appearances in the dataset.


Answer: Opel_Corsa, BMW_118d_DPF, Peugeot_206,BMW_320d_touring and Opel_Corsa_1.2_16V are the first 5 cars that has the highest numbers of appearances in the dataset. 
```{r}
NameFreq <- carsDF %>%
  group_by(name) %>%
  summarise(numberofRow=n()) %>%
  arrange(desc(numberofRow))
head(NameFreq,5)
```


## **New column**
To the dataframe carsDF add a new column called WeekDay_lastseen. WeekDay_lastseen column should store the full name of date from lastSeen Column! 
```{r}
library(lubridate)
Sys.setlocale("LC_TIME", "en_US.UTF-8")
carsDF <- carsDF %>%
  mutate(WeekDay_lastseen=wday(lastSeen, abbr=F, label=T))
head(carsDF$WeekDay_lastseen)
```


## **Number of Saturday cars**
How many cars were last seen on Saturday?
There are 74 cars last seen on Saturday.
```{r}
car_weekday <- carsDF %>%
  group_by(WeekDay_lastseen) %>%
  summarise(numberofWeekdayRow=n())
car_weekday
car_weekday$numberofWeekdayRow[car_weekday$WeekDay_lastseen=="Saturday"]
```


## **Scatterplot**
With carsDF dataframe create a scatterplot in ggplot2 (10)
```{r}
unique(carsDF$gearbox)

scatter_plot <- dplyr::filter(carsDF,gearbox=="manuell" | gearbox=="automatik")

ggplot(data=scatter_plot,mapping = aes(x=price,y=powerPS))+
   geom_point()+
  labs(title="Scatter Plot") +
  scale_x_log10() +
  scale_y_log10() +
  geom_density_2d() +
  facet_wrap(~ gearbox)
```



